### 2014-10-29 src/docs/talks/paris-dauphine-2014/stan-overview.tex
added intro slides back in

>>>>

\sld{Why Stan?}
%
\begin{itemize}
\item {\slshape\bfseries Application}: Fit rich Bayesian statistical models
\end{itemize}
%
\vspace*{2pt}
\begin{itemize}
\item {\slshape Problem}: Gibbs and Metropolis too slow (diffusive) 
\item {\slshape Solution}: Hamiltonian Monte Carlo (flow)
%
\vspace*{8pt}
\item {\slshape Problem}:  Interpreters slow and unscalable
\item {\slshape Solution}: Compiled to C++
%
\vspace*{8pt}
\item {\slshape Problem}:  Need gradients of log posterior for HMC
\item {\slshape Solution}: Reverse-mode algorithmic differentation
\end{itemize}

\sld{Why? (cont.)}
%
\begin{itemize}
\item {\slshape Problem}:  Existing algo-diff slow, limited, unextensible
\item {\slshape Solution}: Our own algo-diff
%
\vspace*{8pt}
\item {\slshape Problem}:  Algo-diff requires functions templated on
  all args
\item {\slshape Solution}: Our own density library, Eigen linear
 algebra
%
\vspace*{8pt}
\item {\slshape Problem}:  Need unconstrained parameters for HMC
\item {\slshape Solution}: Variable transforms w. Jacobian determinants
%
\end{itemize}

\sld{Why? (cont.)}
%
\begin{itemize}
\item {\slshape Problem}:  Need ease of use of BUGS
\item {\slshape Solution}: Compile a domain-specific language
%
\vspace*{8pt}
\item {\slshape Problem}:  Pure directed graphical language inflexible
\item {\slshape Solution}: Imperative probabilistic programming language

\vspace*{8pt}
\item {\slshape Problem}:  Need to tune parameters for HMC
\item {\slshape Solution}: Tune step size and estimate mass matrix
  during warmup;  on-the-fly number of steps (NUTS)
%
\end{itemize}

\sld{Why? (cont.)}
\begin{itemize}
%
\vspace*{8pt}
\item {\slshape Problem}:  Efficient up-to-proportion density calcs
\item {\slshape Solution}: Density template metaprogramming 
%
\vspace*{8pt}
\item {\slshape Problem}:  Limited error checking, recovery
\item {\slshape Solution}: Static model typing, informative exceptions
%
\vspace*{8pt}
\item {\slshape Problem}:  Poor boundary behavior
\item {\slshape Solution}: Calculate limits (e.g. $\lim_{x \rightarrow
    0} x \log x$)
%
\end{itemize}

\sld{Why? (continued)}
\begin{itemize}
\item {\slshape Problem}:  Nobody knows everything
\item {\slshape Solution}: Expand project team with specialists
\vspace*{8pt}
\item {\slshape Problem}:  Nobody knows everything
\item {\slshape Solution}: Expand project team with specialists
\vspace*{8pt}
\item {\slshape Problem}:  Heterogeneous user base
\item {\slshape Solution}: More interfaces (R, Python, MATLAB, Julia),
  domain-specific examples, tutorials

\end{itemize}

\sld{Why? (continued)}
\begin{itemize}
\item {\slshape Problem}:  Expanding project team and interfaces
\item {\slshape Solution}: GitHub process: branch, pull request, code
  review, continuous integration
%
\vspace*{8pt}
\item {\slshape Problem}:  Restrictive licensing (e.g., closed, GPL, etc.)
\item {\slshape Solution}: Open-source, BSD license
\end{itemize}
